---
title: "Demo"
author: "Connie Zhang"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    theme: united
vignette: >
  %\VignetteIndexEntry{Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is a demo vignette for my package. We will generate an adjacency matrix...  

The following lines of code...

```{r}
library(UWBiost561)
set.seed(561)
```

# Generating Partial Cliques

We will generate two partial cliques: one partial clique with 10 nodes, 50% of which are in the clique, and the clique has an edge density of 50%, and another partial clique with 10 nodes, 30% of which are in the clique, and the clique has an edge density of 70%.

```{r}
partial_clique1 <- generate_partial_clique(10,0.5,0.5)$adj_mat
partial_clique2 <- generate_partial_clique(10,0.3,0.2)$adj_mat
partial_clique1
partial_clique2
```

# Reordering Partial Cliques

We will then reorder the partial cliques based on a random permutation of the nodes.

```{r}
order1 <- sample(10,10,replace=FALSE)
order2 <- sample(10,10,replace=FALSE)
reordered_clique1 <- reorder_clique(partial_clique1, order1)
reordered_clique2 <- reorder_clique(partial_clique2, order2)
reordered_clique1
reordered_clique2
```

# Computing Maximal Partial Cliques

Now we compute the maximal partial cliques in the two partial cliques.

```{r}
maximal_clique1 <- compute_maximal_partial_clique(partial_clique1, 0.5)$clique_idx
maximal_clique2 <- compute_maximal_partial_clique(partial_clique2, 0.5)$clique_idx
maximal_clique1
maximal_clique2
```



